---
title: "3C Sustainable Route"
author: "Sandro Araújo, Rosa Félix, Ricardo Ferreira"
date: "2024-08-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      include = FALSE,
                      warning = FALSE,
                      message = FALSE)
```

```{r packages, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(sf)
library(tmap)
tmap_mode("view")
```


## 3C Project

Summary

## A sustainable route connecting the 4 Campi

Only sustainable transportation, such as walking, cycling, train, and ferry.

Destination spots with available amenities, including food, camping, and accomodation.

Route profile for easy ride (not superman/woman).

Stages from 35 to 100 km long.

### Madrid - Coimbra

This connection is a challenge.
The train line which offers the best connection in Portugal is suspended for improvements until 2026.

```{r madcoi}
MADCOI = st_read("route/MAD-COI.gpkg", fid_column_name = "fid", quiet = TRUE)
```

## Madrid - Montpellier

```{r madmon}
MADMON = st_read("route/MAD-MON.gpkg", fid_column_name = "fid", quiet = TRUE)

MADMON1 = MADMON |>
  filter(fid %in% c(2, 3, 5, 6, 7, 13, 14, 15)) |>
  # mutate(lty = ifelse(Mode == "Train", "dashed", "solid"))
  mutate(Stage = as.character(c(0:7)))

MADMON2 = MADMON |> 
   filter(fid %in% c(2,8,12,7,13,14,15)) |> 
   mutate(Stage = as.character(c(0,3,1,2,4,5,6))) |> 
   arrange("Stage")
```


### Option 1

```{r map21, echo=FALSE, include=TRUE}
tm_shape(MADMON1 |> filter(Mode != "Bike")) +
  tm_lines(
    lwd = 2,
    lty = "dashed",
    col = "grey20",
    # lty = "lty", # cannot make this work
    # col = "Stage"
    id = "End",
    popup.vars = c(
      "Stage",
      "Start",
      "End",
      "Distance",
      "Mode",
      "Travel",
      "Acomodation",
      "Notes"
    )
  ) +
  tm_text("Distance") +
  tm_shape(MADMON1 |> filter(Mode == "Bike")) +
  tm_lines(
    lwd = 4,
    # lty = "lty", # cannot make this work
    col = "Stage",
    id = "End",
    popup.vars = c(
      "Stage",
      "Start",
      "End",
      "Distance",
      "Gain",
      "Duration",
      "EuroVelo",
      # "Mode",
      # "Travel",
      "Acomodation",
      "Notes"
    )
  ) +
  tm_text("Distance") +
  tm_title("MAD-MON, option 1")

```

#### Statistics

*   Total Gain = `r sum(MADMON1$Gain, na.rm=T)`
*   Total Days = `r table(MADMON1$Mode)[[1]]`
*   Days on Bike = 6
*   Accommodation cost = `r sum(MADMON1$Acomodation, na.rm=T)`

#### Pros and Cons

* 1 day longer
* shorter stages

<!-- https://www.mapchannels.com/routemaps3/map.htm?title=teste%20rota%201 -->

### Option 2


```{r map22, echo=FALSE}
tm_shape(MADMON2 |> filter(Mode != "Bike")) +
  tm_lines(
    lwd = 2,
    lty = "dashed",
    col = "grey20",
    # lty = "lty", # cannot make this work
    # col = "Stage"
    id = "End",
    popup.vars = c(
      "Stage",
      "Start",
      "End",
      "Distance",
      "Mode",
      "Travel",
      "Acomodation",
      "Notes"
    )
  ) +
  tm_text("Distance") +
  tm_shape(MADMON2 |> filter(Mode == "Bike")) +
  tm_lines(
    lwd = 4,
    # lty = "lty", # cannot make this work
    col = "Stage",
    id = "End",
    popup.vars = c(
      "Stage",
      "Start",
      "End",
      "Distance",
      "Gain",
      "Duration",
      "EuroVelo",
      # "Mode",
      # "Travel",
      "Acomodation",
      "Notes"
    )
  ) +
  tm_text("Distance") +
  tm_title("MAD-MON, option 2")

```



## Montpellier - Turin

```{r montur}
MONTUR = st_read("route/MON-TUR.gpkg", fid_column_name = "fid", quiet = TRUE)
```

